Classical Mechanics Assignment 3

### 2015-10-16 classical/latex/a3.tex
daddsf

\section{Problem Three}
\textbf{A badly designed rocket has an initial mass of 7x104 kg and on firing it burns its fuel at a rate of 250 kg/s. The exhaust speed is 2500 m/s. If the rocket has a vertical ascent from rest on the Earth, how long after the rocket engines fire will the rocket lift off?} \\ \\
\vspace{1 cm}

@@

>>>>

\section{Problem Three}
\textbf{A badly designed rocket has an initial mass of 7x104 kg and on firing it burns its fuel at a rate of 250 kg/s. The exhaust speed is 2500 m/s. If the rocket has a vertical ascent from rest on the Earth, how long after the rocket engines fire will the rocket lift off?} \\ \\
%\vspace{1 cm}
Here $m$ is the mass of the rocket, $m_0$ is the initial mass of the rocket, $dm$ is the mass of the particles being ejected per timestep $dt$, $v$ is the rocket's velocity, $v_{ex}$ is the exhaust's velocity and $dv$ is the rocket's change in velocity per timestep $dt$.

\begin{align*}
  m\ddot{y} &= -mg - \dot{m}v_{ex}\\
  m\ddot{y} &= -mg + |\dot{m}|v_{ex}\\
  \ddot{y} &= -g + \frac{|\dot{m}|}{m}v_{ex}\\
  \frac{dv}{dt} &= -g + \frac{|\dot{m}|}{m}v_{ex}\\
  dv &= \Big(-g + \frac{|\dot{m}|}{m}v_{ex}\Big)dt\\
  dv &= \Big(-g + \frac{|\dot{m}|}{m}v_{ex}\Big)\Big(-\frac{dm}{|\dot{m}|}\Big)\\
  dv &= \Big(\frac{g}{|\dot{m}|} - \frac{v_{ex}}{m}\Big)dm\\
  \int_0^v dv &= \int_{m_0}^m \Big(\frac{g}{|\dot{m}|} - \frac{v_{ex}}{m}\Big)dm\\
  v &= \frac{g}{|\dot{m}|}\big(m-m_0) - v_{ex}\log(\frac{m}{m_0})\\
  v &= \frac{-g}{|\dot{m}|}\big(m_0-m) + v_{ex}\log(\frac{m_0}{m})\\  
\end{align*}

This is an incorrect equation in that it predicts negative velocity when the gravitational force is stronger than the exhaust force. This is because we didn't include the normal force term in our initial enumeration of the forces acting on the rocket. The second thrust exceeds gravity the normal force disappears and our equation becomes correct. This coincides with the time the rocket lifts off:\\

\begin{align*}
  0 &= \frac{-g}{|\dot{m}|}\big(m_0-m) + v_{ex}\log(\frac{m_0}{m})\\
  \frac{-g}{|\dot{m}|}\big(m_0-m) &= v_{ex}\log(\frac{m_0}{m})\\
\end{align*}

We plug in our values:\\

\begin{align*}
  \Big(\frac{-9.8\frac{m}{s^2}}{250\frac{kg}{s}}\Big)\big(7*10^4kg - 250\frac{kg}{s}t) &= \big(2500\frac{m}{s}\big)\log\Big(\frac{7*10^{4}kg}{250\frac{kg}{s}t}\Big)\\
  \Big(1.568^{e-5}\frac{1}{kg}\Big)\big(7e4 kg - 250*t kg) &= \log\Big(\frac{7e4}{250*t}\Big)\\
  1.0976 - 0.00392t &= \log\Big(\frac{280}{t}\Big)\\
  -4.5372 - 0.00392t &= -\log(t)\\
  t &= 231.752
\end{align*}

@@

Using conservation of momentum and conservation of energy, we can form two equations:\\

\begin{align*}
  MV &= \frac{1}{2}MV_1 + \frac{1}{6}MV_2\\
  3MV^2 &= \frac{1}{4}MV_1^2 + \frac{1}{12}MV_2^2\\
\end{align*}

We have two equations and two unknowns. We can solve for $V_1$ and $V_2$:\\

\begin{align*}
  V &= \frac{1}{2}V_1 + \frac{1}{6}V_2\\
  V^2 &= \frac{1}{4}V_1^2 + \frac{1}{36}V_2^2 + \frac{1}{6}V_1V_2\\
  3V^2 &= \frac{1}{4}V_1^2 + \frac{1}{12}V_2^2\\
  V^2 &= \frac{1}{12}V_1^2 + \frac{1}{36}V_2^2\\
  \frac{1}{4}V_1^2 + \frac{1}{36}V_2^2 + \frac{1}{6}V_1V_2 &= \frac{1}{12}V_1^2 + \frac{1}{36}V_2^2\\
  \frac{1}{6}V_1^2 + \frac{1}{6}V_1V_2 &= 0\\
  V_1 &= -V_2\\
\end{align*}

We plug this back into our first equation...\\

\begin{align*}
  V &= \frac{1}{2}V_1 + -\frac{1}{6}V_1\\
  V &= \frac{1}{3}V_1\\
  V_1 &= 3V\\
  V_2 &= -3V\\
\end{align*}
